
---
title: "Homework 1 (tm2925)"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## HW 1: Problem 1
Create a data frame with a random sample of 10 with a uniform distribution, logical vector witih elements greater than 2, character vector and factor vector.

```{r data frame for problem 1}
hw1_q1_df = tibble(
  samp = runif(10, 0, 5),
  vec_logical = samp > 2,
  vec_character = c('a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j'),
  vec_factor = factor(c(vec_character))
)

##taking the mean of each variable
mean(hw1_q1_df$vec_logical)
mean(hw1_q1_df$vec_character)
mean(hw1_q1_df$vec_factor)
```
**What works and what doesn’t? Why?**
_The mean of a logical vector works, the mean of character or factor vectors do not work._

```{r coerce as numeric, eval=FALSE}
##coercing vectors
as.numeric(hw1_q1_df$vec_logical)
as.numeric(hw1_q1_df$vec_character)
as.numeric(hw1_q1_df$vec_factor)
```

```{r convert}
##convert your character variable from character to factor to numeric; convert your factor variable from factor to character to numeric
as.numeric(as.factor(hw1_q1_df$vec_character))
as.numeric(as.character(hw1_q1_df$vec_factor))
```

**What happens?**
_Character vectors are not able to be coerced into numeric vectors. This causes a warning message and NA._

